Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-11-19 ba_resubmit/response_to_reviewers.tex

%\textcolor{blue}{This is a good point. I wonder if we could perhaps remove a lot of the notation by Sadinle/FS and redo the notation such that the ordering wouldn't matter. Also, it would make more sense to follow with the examples that we use in the paper -- NLTCS and EL Salvador so that we don't confuse the reader. Let's talk through this point as this would be a big change. If the algorithm and notation can be easily modified to reverse things, this would be a nice addition to the paper and improvement over Sadinle.}

\pointRaised{R1}{%
5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript. This consideration of missing fields and the accompanying discussion of missing data mechanism should be included if such data is to be considered for this manuscript. This comment also applies to the $Gamma^{obs}$ notation in the following subsection.
}

\reply{%
	Our missingness assumptions and the definition of $I_{obs}()$ have been added right before Section 2.1.
}

@@

\reply{%
	The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior. The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$. We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$. We have updated our explanation to make this more clear.
}

\pointRaised{R1}{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-$\pi$).
}
\reply{%
	This has been corrected.
}

\pointRaised{R1}{%
Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
}

\reply{%
	We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have updated additional parts of the paper with this revision. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.
}

@@

\pointRaised{R1}{%
	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?
}

\reply{%
	In the caption under Figure 3, we explain that we have 100 pairs datasets for each level of overlap and error. Thus the table represents results from 900 pairs of datasets. We have added this description in the body of the paper as well for clarity. 
}

\pointRaised{R1}{%
Section 5.3 simulation – RR results are not presented.
}

\reply{%
	We have updated the Figure with these results. Since it is presented alongside NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results, instead of rejection rate (RR) results. Additionally, this analysis was moved to the appendix so the main body more clearly focuses on our contributions.
}

### 2022-11-19 ba_resubmit/kundinger_resubmit.tex
final revisions

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
\textcolor{red}{Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages.} However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two datafiles. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two datafiles that contain duplications across, but not within, the respective datafiles.

Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. Additionally, some methods model records directly \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a datafile, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on datafiles with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger datafiles than previous Bayesian Fellegi Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each datafile, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller datafile.

@@

Consider two datafiles $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in X_1$" rather than the more compact $x_{1i}$ \textcolor{red}{for ease of readability}. Under bipartite matching, the set of matches across datafiles can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets. We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For numerical features, we can use absolute difference between the two feature values. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

\textcolor{red}{In the construction of comparison vectors, we note it is common to encounter missing information in datasets used for record linkage. Whether information is missing in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, by \cite{LittleRubin2002}). With the MAR, we can marginalize over the missing data, and do all computation simply using the observed data. To notate this, we use the indicator $I_{obs}(\cdot)$, which takes on the value 1 when the argument inside the parentheses is observed, and 0 otherwise.}

@@

The seminal \cite{fellegi_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. These allow us to model the comparison data using a mixture model, where
\begin{align}
	\label{eqn:fs_model}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
\end{align}
Here $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match. Such independent classifications often leads to matchings that violate one-on-one matching assumptions, which prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi Sunter weights among matched pairs.

@@

%The Fellegi Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

%The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
%\end{align}
%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
%
%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
%
%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
%
%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast Beta prior as follows:
\begin{align}
	\label{eqn:fast_beta_prior}
	Z_j | \pi &= \begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret equation \ref{eqn:fast_beta_prior} as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$

\textcolor{red}{Previously, \cite{wortman2019} proposed a flat prior over the labeling of $\bm{Z}$, meaning that each value $\{1, \ldots, n_1, n_1 +j\}$ was considered equally likely for $Z_j$. This however amounts to a prior probability of $\frac{n_1}{n_1 + 1}$ that record $j$ has a match in $\bm{X}_1$, and thus tended to over match, leading to poor precision when used in practice. Indeed we find that the flat prior is equivalent to a special case of the fast Beta prior where $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.}

@@

	\section*{Appendix}
	\label{sec:appendix}

>>>>

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

In many data analysis tasks, analysts seek to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified each pair as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

@@

Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}). With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{align}
	\label{eqn:fs_model}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \notag
\end{align}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

%\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
%\end{align}
%This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match.

Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters. However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results. To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching. \cite{sadinle_bayesian_2017} later showed that this method could be thought of as a maximum likelihood estimate.

%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.

@@

%The Fellegi-Sunter method then uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i', j)$ are declared as links, even though such a matching is not allowed by assumption.

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

%The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
%\end{align}
%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
%
%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
%
%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
%
%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast Beta prior as follows. For each $Z_j$, we have
\begin{align}
	\label{eqn:fast_beta_prior}
	p(Z_j = z_j| \pi) &= \begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$. This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$. In our simulation studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.

@@

	\section{Appendix}
	\label{sec:appendix}
